---
title: "Exercise - dplyr"
output: html_notebook
---

> Run following command in R console to link to `dplyr` library:

```{r}
library(dplyr)
```


### Piping
> Now let's try to calculate simple thing using multiple encapsulated functions.

```{r}
abs(min(c(1, 4, 2, 8, -5)))
```
> Can you understand what this command does? Is the result what you've expected?  
> With `dplyr` we can code the same thing using *piping*

```{r}
c(1, 4, 2, 8, -5) %>% min %>% abs
```
> Result is the same!


### 6 basic `dplyr` verbs
> In this exercise let's practice `dplyr` and *piping* on simple dataset. Firstly let's `read` some data in:  

```{r}
dt_KPI <- read.csv("./Data/lesson2_KPI.csv")
```

> We can select top 6 rows using `head` function. 

```{r}
dt_KPI %>% head
```

> OK, now let's create additonal column *Combined* by `mutate` and showing first 6 rows right away.  

```{r}
dt_KPI %>% 
  mutate(Combined = Losses + Expenses) %>% head
```

> What if we want to see only *Region* and *Losses*? Let's `select` it to output:

```{r}
dt_KPI %>% 
  select(Region, Losses) %>% head
```

> We can also `filter` data using condition on one (or more) columns, for instance returning *Criminals* only:

```{r}
dt_KPI %>% 
  filter(Business == "Criminals") %>% head
```

> In some cases it is good to `arrange` the data to have a better picture about it (eg order data by *Premium*):

```{r}
dt_KPI %>% 
  arrange(Losses, desc(Premium)) %>% head
```

> Or we may want to `summarise` data to see some basic statistic about it, like average losses.

```{r}
dt_KPI %>% 
  summarise(Loss_avg = mean(Losses))
```

> We can see the summary split by one (or more) of columns - in this case average losses per Region

```{r}
dt_KPI %>% 
  group_by(Region) %>% 
  summarise(Loss_avg = mean(Losses)) %>% head
```

> And at last... remember that the best way to master any topic is to practice and be curious!