---
title: "Data preparation for lesson 2 (KPI's, searching for bad looking portfolios)"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
    toc: yes
---

# Setup

## Load libraries

```{r}
# libraries
# install.packages("tidyr")
library(dplyr)
library(ggplot2)
library(tidyr)
```

## Load data

```{r}
dt_KPI_clean <- readRDS("/cloud/project/Data/lesson2_KPI_clean.rds")
```

# Inspecting basic insurance metrics
We talked about three basic insurance metrics - Premium, Losses and Expenses. We can sum these up to see the overall statistics of our dataset (insurance company).

```{r}
dt_KPI_clean %>%
  summarize(
    Premium = sum(Premium),
    Losses = sum(Losses),
    Expenses = sum(Expenses)
  )
```

# Visualization of Raw Data 
*What makes sense to visualize (or compare using some visualization) in terms of KPIs?*
Reflection - take a moment and think about what makes sense to look into.

## Data preparation for visualization
### Which Unit has collected the most Premium? 

```{r}
dt_KPI_clean %>%
  group_by(Unit) %>%
  summarize(Premium = sum(Premium)) %>%
  arrange(desc(Premium))
```

### Which Unit has the lowest Expenses? 
```{r}
dt_KPI_clean %>%
  group_by(Unit) %>%
  summarize(Expenses = sum(Expenses)) %>%
  arrange(Expenses)
```

### Which Business from Unit found in the previous task had the highest Expenses? 
```{r}
dt_KPI_clean %>%
  filter(Unit == "Unit8") %>%
  group_by(Business) %>%
  summarise(Expenses = sum(Expenses)) %>%
  arrange(desc(Expenses))
```

## Basic Visualization

Visualize your findings on simple bar charts - ggplot2::geom_col()

### Which Unit has collected the most Premium? 
```{r}
dt_KPI_clean %>%
  group_by(Unit) %>%
  summarize(
    Premium = sum(Premium),
    Expenses = sum(Expenses),
    Losses = sum(Losses)
  ) %>%
  ggplot(aes(
    x = reorder(Unit, -Premium),
    y = Premium,
    group = 1
  )) +
  # geom_line(), geom_point(), geom_hist(), ...
  geom_col()
# +
# geom_smooth() +
# geom_point()
```

### Which Unit has the lowest Expenses? 
```{r}
dt_KPI_clean %>%
  group_by(Unit) %>%
  summarize(
    Premium = sum(Premium),
    Expenses = sum(Expenses),
    Losses = sum(Losses)
  ) %>%
  ggplot(aes(
    x = reorder(Unit, Expenses), y = Expenses,
    group = 1
  )) +
  geom_point() +
  geom_line()
```

### Which Business from Unit found in the previous task had the highest Expenses? 
```{r}
dt_KPI_clean %>%
  filter(Unit == "Unit5") %>%
  group_by(Business) %>%
  summarize(
    Premium = sum(Premium),
    Expenses = sum(Expenses),
    Losses = sum(Losses)
  ) %>%
  ggplot(aes(x = reorder(Business, -Expenses), y = Expenses)) +
  geom_col()
```

### Bonus - Show all Costs Insurance Company has as a stacked bar for Expenses and Losses grouped by Units

Summarize Expenses and Losses at Unit level:

```{r}
dt_by_unit <- dt_KPI_clean %>%
  group_by(Unit) %>%
  summarise(
    Expenses = sum(Expenses),
    Losses = sum(Losses)
  )

dt_by_unit
```

Pivot summarized data to long form: (if you are getting an error, you might need to install the tidyr package using `install.packages("tidyr")`)

```{r}
dt_by_unit_longer <- dt_by_unit %>%
  tidyr::pivot_longer(-Unit, names_to = "Type", values_to = "Value")

dt_by_unit_longer
```

Visualise as stacked bar chart:

```{r}
dt_by_unit_longer %>%
  ggplot(aes(x = Unit, y = Value, fill = Type)) +
  geom_col()
```
We can use the argument `position` of `geom_col` function to arrange the bars to be side by side.

```{r}
dt_by_unit_longer %>%
  ggplot(aes(x = Unit, y = Value, fill = Type)) +
  geom_col(position = "dodge")
```

We can also use a different value for this argument to normalize values to [0.0, 1.0] within each x-axis category (Unit).

```{r}
dt_by_unit_longer %>%
  ggplot(aes(x = Unit, y = Value, fill = Type)) +
  geom_col(position = "fill")
```

Reflection - take a moment and think which is the best and the worst Unit performer. Feel free to do more data exploration.
