---
title: "Enhancing GLM with GBM and Model Evaluation"
output:
  html_document:
    df_print: paged
---

*Before running chunks within this markdown file, run all chunks in 01_GBM_Basic.Rmd and 02_GBM_Tuning.Rmd.*

```{r load setup}
# Load all required files. function, packages
libs <- c("Ckmeans.1d.dp")
lapply(libs, function(lib) {
  if (!lib %in% installed.packages()) install.packages(lib)
})

# Load the required packages
library(dplyr)
library(xgboost)
library(ggplot2)
library(caret)
library(plotly)
library(tidyr)

# Source support functions
functions_path <- file.path("/cloud/project/Support/functions")
functions_files <- list.files(
  functions_path, 
  pattern = "\\.R$"
)

invisible(lapply(file.path(functions_path, functions_files), source))

# Load all necesarry objects from previuos reports
data_part_03 <- readRDS(file = "/cloud/project/Data/data_part_03.rds")

xgb_final <- data_part_03$xgb_final
test <- data_part_03$test
oneway_plots <- data_part_03$oneway_plots
```

As mentioned in the GBM motivation, we can *use GBM to improve GLM*. In the previous parts, we ran a basic GBM, tuned the GBM, and found the model with the best performance. We also learned how to interpret the model. Now, we will use this knowledge from GBM to enhance the performance of our GLM.

The key idea here is to leverage the insights gained from the GBM to refine our GLM. Specifically, we will select only the most important variables identified by the GBM, adjust them using PDP and evaluate the performance of the GLM model. This approach helps us create a more efficient and potentially more accurate GLM by focusing on the most impactful predictors.

In this script, we will then compare the GLM model created using insights from the GBM with the GLM models that were created in Lesson 5. We will also evaluate the impact of adding or removing variables and their feature engineering on the model's performance. This comparison will help us understand how the enhancements based on GBM insights affect the GLM's accuracy and reliability and why feature engineering and data preparation are important steps in the modeling process.

Let's look at the tuned GBM model. We can see that there are still some overestimated or underestimated areas on the lift curve, but overall, the model fits well. (However, there is still room for improvement, and you can try to improve the model at home by adjusting the hyperparameters or feature engineering.)

```{r lift chart}
# Overall model performance
lift_curve(
  data = test,
  target_variable = "Burning_Cost",
  weights_input = "Time_Exposure",
  pred_fun = calculate_predictions,
  is_apply_onehot = TRUE,
  object = xgb_final
)
```

The feature importance chart shows which variables are the most important in our model. The most important variables are BonusMalus, Sum_insured, Constructed_year, Capacity, D_age, etc. These are the variables we will include in our final GLM model.

```{r feature importance}
# The most important variables
xgb_fi <- xgb.importance(model = xgb_final)
xgb.ggplot.importance(xgb_fi)
```

Let's also look at the trends of these important variables. For example, on the BonusMalus chart, we can see that as the BonusMalus level increase, the Burning Costs decrease.

We can see that the most important variables are numeric. Now, we will adjust them, band or cap them by PDP, and change the variable type of some variables to categorical ones to improve the model.

```{r oneway}
# One way plot with PDP 

# important_vars <- xgb_fi$Feature[1:7]
# 
# for (var in important_vars) {
#    calculate_oneway(
#     data = test,
#     model = xgb_final,
#     var = var,
#     pred_fun = calculate_predictions,
#     is_apply_onehot = TRUE,
#     target_name = "Burning_Cost",
#     weight_name = "Time_Exposure"
#   )
# }

oneway_plots
```

Firstly load the data and models from Lesson 5, adjust the most important variables on these data, and create a new GLM model using the adjusted most important variables.

```{r}
# Load models and data from lesson 5
glm_models <- readRDS(file = "/cloud/project/Data/glm_models.rds")

train <- glm_models$train
train_actual <- train$Burning_Cost
train_weight <- train$Time_Exposure
valid <- glm_models$valid
```

```{r}
# Adjust the most important variables by PDP 

# Train 
train <- train %>% 
  mutate(
    # Capacity PDP
    #  0-1300 flat, 1300-2000, 2000-3000, 3000 - INF
    Capacity_banded = cut(
      Capacity,
      breaks = c(-Inf, 1300, 2000, 3000, Inf),
      dig.lab = 6
    ),
    # D_age
    #  decreasing up to 40, then increasing and after 70 flat
    D_age_banded = cut(
      D_age,
      breaks = c(-Inf, seq(20, 70, 10), Inf),
      dig.lab = 6
    ),
    # Sum_insured increasing trend up to 15000 - than flat
    Sum_insured_capped = ifelse(Sum_insured > 15000, 15000, Sum_insured),
    # Small amount of levels - categorical variables
    BonusMalus_capped = as.factor(ifelse(BonusMalus < 0, 0, BonusMalus)),
    Time_on_book_capped = as.factor(ifelse(Time_on_book > 20, 20, Time_on_book))
  )

# Valid 
valid <- valid %>% 
  mutate(
    # Capacity PDP
    #  0-1300 flat, 1300-2000, 2000-3000, 3000 - INF
    Capacity_banded = cut(
      Capacity,
      breaks = c(-Inf, 0, 1300, 2000, 3000, Inf),
      dig.lab = 6
    ),
    # D_age
    #  decreasing up to 40, then increasing and after 70 flat
    D_age_banded = cut(
      D_age,
      breaks = c(-Inf, seq(20, 70, 10), Inf),
      dig.lab = 6
    ),
    # Sum_insured increasing trend up to 20000 - than flat
    Sum_insured_capped = ifelse(Sum_insured > 20000, 20000, Sum_insured),
    # Small amount of levels - categorical variables
    BonusMalus_capped = as.factor(ifelse(BonusMalus < 0, 0, BonusMalus)),
    Time_on_book_capped = as.factor(ifelse(Time_on_book > 20, 20, Time_on_book))
  )
```

Let's create a GLM model using insights from GBM and include the variables identified as most important. However, since we identified that Veh_type2 is correlated with Construct_year and was an important variable in Lesson 5 models, we will use the Veh_type2 variable instead of Capacity.

Let's fit two models: the first with the 7 most important variables, and the second with the 3 most important variables.

```{r}
# Enhancing GLM with GBM
# 7 most important variables by GBM
final_glm1 <- glm(
  data = train,
  formula = Burning_Cost ~ BonusMalus_capped + Sum_insured_capped + Veh_type2_merged + D_age_banded + Time_on_book_capped + Year,
  family = statmod::tweedie(var.power = 1.5, link.power = 0),
  weights = Time_Exposure
)

# 3 most important variables by GBM
final_glm2 <- glm(
  data = train,
  formula = Burning_Cost ~ BonusMalus_capped + Sum_insured_capped + Veh_type2_merged,
  family = statmod::tweedie(var.power = 1.5, link.power = 0),
  weights = Time_Exposure
)
```

Now, let's compare the models from Lesson 5 with our enhanced GLM models using insights from GBM to see if the new models are better than the previous ones.

```{r}
# Create list of models
model_list <- list(
  "model1 \n add \n Veh_Type2" = glm_models$model1,
  "model2 \n add \n Construct_Year" = glm_models$model2,
  "model3 \n merge \n Veh_Type2" = glm_models$model3,
  "model4 \n cap \n Construct_Year" = glm_models$model4,
  "model5 \n GLM with GBM \n 7 vars" = final_glm1,
  "model6 \n GLM with GBM \n 3 vars" = final_glm2
)

# Compare models RMSE
compare_models(model_list)
```

We'll create a small visualization on top of the table with metrics.

```{r}
# Plot model comparison
compare_models(model_list) %>%
  ggplot(aes(x = model_nm, y = train_rmse, group = 1)) +
  geom_line() +
  geom_point() +
  ggtitle("Training RMSE")

compare_models(model_list) %>%
  ggplot(aes(x = model_nm, y = valid_rmse, group = 1)) +
  geom_line() +
  geom_point() +
  ggtitle("Validation RMSE")
```
