---
title: "Model_Evaluation"
output:
  html_document:
    df_print: paged
---

Let's come back to the validation dataset and use it to evaluate the impact that adding/removing variables and their feature engineering had on the model performance on both training and validation data.

We'll create a list of all the models we had (starting with model1).

```{r}
model_list_prep <- readRDS(file = "Data/model_list_prep.rds")
train <- model_list_prep$train
valid <- model_list_prep$valid
train_actual <- model_list_prep$train_actual
train_weight <- model_list_prep$train_weight

model_list <- list(
  "model1 \n add Veh_Type2" = model_list_prep$model1,
  "model2 \n add Construct_Year" = model_list_prep$model2,
  "model3 \n merge Veh_Type2" = model_list_prep$model3,
  "model4 \n cap Construct_Year" = model_list_prep$model4
)
```

We'll now use a function that is already prepared in `Support` folder to compute the model performance using RMSE metric.

```{r}
source("Support/compare_models.R")
source("Support/rmse_group.R")

compare_models(model_list)
```

We'll create a small visualisation on top of the table with metrics

```{r}
library(ggplot2)
library(tidyr)

compare_models(model_list) %>%
  ggplot(aes(x = model_nm, y = train_rmse, group = 1)) +
  geom_line() +
  geom_point() +
  ggtitle("Training RMSE")

compare_models(model_list) %>%
  ggplot(aes(x = model_nm, y = valid_rmse, group = 1)) +
  geom_line() +
  geom_point() +
  ggtitle("Validation RMSE")
```

