---
title: "Final Model Interpretation"
output:
  html_document:
    df_print: paged
---

When you are satisfied with the variables that you used and the adjustments you applied, it is now time to use both training and validation dataset together to fit the final model and obtain the most accurate coefficients / multipliers that will be used for pricing use case. 

```{r set up}
# Load the required packages
library(tidyverse)
library(plotly)
options(scipen=999)

# Source support functions
functions_path <- file.path("/cloud/project/Support/functions")
functions_files <- list.files(
  functions_path, 
  pattern = "\\.R$"
)

invisible(lapply(file.path(functions_path, functions_files), source))

# Load models and data from lesson 5
glm_models <- readRDS(file = "/cloud/project/Data/glm_models.rds")
train <- glm_models$train
valid <- glm_models$valid
```


```{r data prep}
# Bind train and valid data
final_data <- rbind(train, valid)

# Adjust predictors
final_data <- final_data %>% 
  mutate(
    Sum_insured_capped = round(ifelse(Sum_insured > 15000, 15000, Sum_insured), -3),
    BonusMalus_capped = as.factor(ifelse(BonusMalus < 0, 0, BonusMalus))
  )
```

```{r fit final model}
final_glm <- glm(
  data = final_data,
  formula = Burning_Cost ~ BonusMalus_capped + Sum_insured_capped + Veh_type2_merged,
  family = statmod::tweedie(var.power = 1.5, link.power = 0),
  weights = Time_Exposure
)
```

```{r explore model}
summary(final_glm)
```

```{r one-way analysis}
fitted_vars <- get_vars(final_glm)

for (var_nm in fitted_vars) {
  emb_chart_plotly(
    model = final_glm,
    x_var = var_nm
  ) %>%
    show()
}
```

Applying the inverse log link function will result in multipliers, that can now be used in pricing context.

```{r}
# Print estimated coefficients (betas) 
round(summary(final_glm)$coef[, 1], 5)
# Print multipliers used in pricing context
round(exp(summary(final_glm)$coef[, 1]), 5)
```


An example:
- Base Price - Set by our business partners, our intercept will not be used

- Car Type:
  - Personal car: 1.00 multiplier to the base price
  - Motorbike : 0.91 multiplier to the base price
  - Pickup: 0.98 multipler to the base price
  - Truck: 1.47 multiplier to the base price
  
- Sum Insured:   
  - With numeric variables, the multiplier has to be computed from the original coefficients
  - Set a reference value, e.g. 1000 because it has the most exposure
  - The multiplier for a new value, e.g. sum insured 2000 will be 1.03 computed as follows:
    - exp(beta * value_new) / exp(beta * value_ref)
    - exp(0.00003 * 2000) / exp(0.00003 * 1000)
    
- Final price = Base Price * Car Type multiplier * Bonus Malus multiplier * Sum Insured multiplier

- There can be another set of multipliers that are applied on top of the final price in the pricing context:
  - Profit multiplier (we also need to earn something on top of the losses)
  - Portfolio multiplier (are we planning to globally move the portfolio to higher/lower prices)
  - Underwriter adjustment based on business knowledge and/or customer relationships
  - etc.
  
This pricing structure is then implemented/updated in an IT system that is used by employees of insurance company when setting prices for insurance policies for next year.

```{r}
exp(summary(final_glm)$coef[, 1])

# Vehicle Type
broom::tidy(final_glm) %>%
  filter(stringr::str_detect(term, "Veh_type2")) %>%
  select(term, estimate) %>%
  mutate(multiplier = exp(estimate))

# Bonus Malus
broom::tidy(final_glm) %>%
  filter(stringr::str_detect(term, "BonusMalus")) %>%
  select(term, estimate) %>%
  mutate(multiplier = exp(estimate))

# Sum Insured
beta <- broom::tidy(final_glm) %>%
  filter(stringr::str_detect(term, "Sum_insured_capped")) %>%
  pull(estimate) %>%
  as.numeric()

all_levels <- final_data$Sum_insured_capped  %>%
  unique() %>%
  sort()

# exp(beta * value_new) / exp(beta * value_ref)
data.frame(Sum_insured_capped = all_levels) %>%
  mutate(
    estimate = beta * as.numeric(Sum_insured_capped),
    estimate_ref = beta * 1000,
    multiplier = exp(estimate) / exp(estimate_ref)
  )
```

Short feedback form for this class - https://forms.office.com/e/0pijTdVZSR
