---
title: "Loss Development Intro"
output:
  html_document:
    df_print: paged
---

### Triangles in insurance

Do you know what this is? It is the core of Reserving and in many cases also pricing. This is also how the losses from previous lessons were produced. It is a very simple way to get many answers like:

1)  How long does it take to fully pay the claims?
2)  How fast are they paid?
3)  Is the initial estimate of the future losses conservative or optimistic?
4)  Once we know the above, how much money can we invest and when? (Because we want to earn interest on premium)

There are far more questions and answers available, and there are many methods that can do the job. But for now, let's understand at least the basics.

### But why should we care?

Insurance company aims to protect customers from risks at *most appropriate* price. This means we cannot waste any resources, and we need to make a machine out of it. We pay claims *just-in-time* (not too early and not too late) and we invest money in financial derivatives with appropriate duration.

On top of this there is also someone else asking us to be fair. It is the *regulator*. Insurance industry is heavily regulated and frequently audited to assure the companies are solvent and ready to pay to customers when needed, rather then to its directors, when premium is collected.

So that is why - customers and regulators. This means it truly is important. So now when we know why, let's look at *how*.

### What is a triangle

As losses take sometimes a number of years to be reported and then some more years to be paid, it is very hard to say at the end of the year whether we have "saved" enough money. For this reason, it makes sense to keep history of all claims related to 1 period.

Imagine this simple situation that shows a history of a single policy:

1)  1st December 2016 - insurance contract was signed to protect against car crash injury
2)  20th November 2017 - there was an accident. The person goes to hospital and does not care about calling insurance company
3)  1st January 2018 - person is finally home, but he will not be able to work for unknown time
4)  5th January 2018 - claim is reported. Doctors say, the person may require another surgery in 6 months
5)  1st February 2018 - new symptoms are found (they seem to be coming from the accident) ...
6)  a few years later - the person finally starts working, but is partially disabled

In the database of an insurance company, this policy history could look something like this:

```{r, echo=F}
tibble::tribble(
  ~AY, ~DY, ~Year, ~Amount, ~Note,
  NA, NA, 2016, 0, "Policy Start",
  NA, NA, 2017, 0, "Claim happened",
  NA, NA, 2018, 0, "Claim reported",
  NA, NA, 2018, 10000, "Car",
  NA, NA, 2018, 1000, "Wage compensation for 2017/12",
  NA, NA, 2018, 3000, "Surgery1",
  NA, NA, 2018, 4000, "Surgery2",
  NA, NA, 2018, 12000, "Wage compensation for 2018/01-2018/12",
  NA, NA, 2019, 12000, "Wage compensation for 2019/01-2019/12",
  NA, NA, 2020, 2400, "Partial disability compensation",
  NA, NA, 2021, 2400, "Partial disability compensation",
)
```

*Think about which year each information relates to*, since for the ease of archivation and reporting, the insurance company needs to relate all this history to a single year.

Once we choose this year, we can reflect this information in our database:

```{r, echo=F}
tibble::tribble(
  ~AY, ~DY, ~Year, ~Amount, ~Note,
  NA,   NA,  2016,  0,       "Policy Start",
  2017, 1,   2017,  0,       "Claim happened",
  2017, 2,   2018,  0,       "Claim reported",
  2017, 2,   2018,  10000,   "Car",
  2017, 2,   2018,  1000,    "Wage compensation for 2017/12",
  2017, 2,   2018,  3000,    "Surgery1",
  2017, 2,   2018,  4000,    "Surgery2",
  2017, 2,   2018,  12000,   "Wage compensation for 2018/01-2018/12",
  2017, 3,   2019,  12000,   "Wage compensation for 2019/01-2019/12",
  2017, 4,   2020,  2400,    "Partial disability compensation",
  2017, 5,   2021,  2400,    "Partial disability compensation",
)
```

We can aggregate this information for each combination of AY and DY, then display in a wide (pivoted) format like this:

```{r, echo=F}
tibble::tribble(
  ~AY, ~DY1, ~DY2, ~DY3, ~DY4, ~DY5,
  2017, 0, 30000, 42000, 44400, 46400
)
```

However, since we're an insurance company, *we're interested in looking at the whole portfolio* (multiple policies and therefore multiple claims). Let's assume we had 10 similar claims that originated in 2017 (accident year):

```{r, echo=F}
tibble::tribble(
  ~AY, ~DY1, ~DY2, ~DY3, ~DY4, ~DY5,
  2017, 0, 300000, 420000, 444000, 464000
)
```

Working with a bigger portfolio also means, that *some claims usually happen every year*, not just in 2017. So a *full picture of our portfolio* could look something like this. Note that as we progress in time (AY), we have less and less information on how the losses will look like in the following development years (DY).

This structure is shaped like a triangle, where the rows represent the years of origin (policy, accident or reporting year) and the columns represent how long it took until the information was recognized.

```{r, echo=F}
tibble::tribble(
  ~AY, ~DY1, ~DY2, ~DY3, ~DY4, ~DY5,
  2017, 0, 300000, 420000, 444000, 464000,
  2018, 10000, 320000, 410000, 454000, NA,
  2019, 12000, 350000, 430000, NA, NA,
  2020, 6000, 305000, NA, NA, NA,
  2021, 9000, NA, NA, NA, NA
)
```

### What do we do with this?

Well, we want to take the advantage of the history that we have collected and predict what is likely to happen with the latest year, that we have just got to know about. It is basically a variation of the Linear Regression, where you learn from past patterns and apply them to predict future.

To get what we are talking about, it is worth visualizing this a bit. Use the ChainLadder package and explore the triangle object from example there.
