---
title: "Loss Development Intro"
output:
  html_document:
    df_print: paged
---

### Triangles in insurance

Do you know what this is? It is the core of Reserving and in many cases also pricing. This is also how the losses from previous lessons were produced. It is a very simple way to get many answers like:

1)  How long does it take to fully pay the claims?
2)  How fast are they paid?
3)  Is the initial estimate of the future losses conservative or optimistic?
4)  Once we know the above, how much money can we invest and when? (Because we want to earn interest on premium)

There are far more questions and answers available, and there are many methods that can do the job. But for now, let's understand at least the basics.

### But why should we care?

Insurance company aims to protect customers from risks at *most appropriate* price. This means we cannot waste any resources, and we need to make a machine out of it. We pay claims *just-in-time* (not too early and not too late) and we invest money in financial derivatives with appropriate duration.

On top of this there is also someone else asking us to be fair. It is the *regulator*. Insurance industry is heavily regulated and frequently audited to assure the companies are solvent and ready to pay to customers when needed, rather then to its directors, when premium is collected.

So that is why - customers and regulators. This means it truly is important. So now when we know why, let's look at *how*.

### What is a triangle

As losses take sometimes a number of years to be reported and then some more years to be paid, it is very hard to say at the end of the year whether we have "saved" enough money. For this reason, it makes sense to keep history of all claims related to 1 period.

Imagine this simple situation. **Driver A** holds a PZP policy — mandatory third-party liability insurance that covers damages they may cause to others. **Person B** is someone Driver A will injure in an accident.

1)  1st December 2021 — Driver A signs a PZP contract. This covers bodily injury and property damage they may cause to third parties.
2)  20th November 2022 — Driver A causes an accident, injuring Person B. Person B is rushed to hospital. Driver A reports the accident to their insurer within a few days, as required by law.
3)  December 2022 — While in hospital, Person B's family contacts SKP (the national motor insurance bureau) with Driver A's license plate and identifies the insurer. They notify the insurer of the injury, but Person B is in no condition to formally submit a claim yet.
4)  1st January 2023 — Person B returns home after weeks in hospital, but will not be able to work for an unknown period.
5)  5th January 2023 — Person B formally submits a claim to Driver A's insurer. Doctors say another surgery may be needed in 6 months.
6)  1st February 2023 — new symptoms are found (they seem to be related to the accident) ...
7)  a few years later — Person B finally starts working again, but is partially disabled.

In the database of an insurance company, this policy history could look something like this:

```{r, echo=F}
tibble::tribble(
  ~AY, ~DY, ~Year, ~Amount, ~Note,
  NA, NA, 2021, 0, "Policy Start",
  NA, NA, 2022, 0, "Accident (Driver A reports to insurer)",
  NA, NA, 2023, 0, "Claim submitted by Person B",
  NA, NA, 2023, 10000, "Car",
  NA, NA, 2023, 1000, "Wage compensation for 2022/12",
  NA, NA, 2023, 3000, "Surgery1",
  NA, NA, 2023, 4000, "Surgery2",
  NA, NA, 2023, 12000, "Wage compensation for 2023/01-2023/12",
  NA, NA, 2024, 12000, "Wage compensation for 2024/01-2024/12",
  NA, NA, 2025, 2400, "Partial disability compensation",
  NA, NA, 2026, 2400, "Partial disability compensation",
)
```

*Think about which year each information relates to*, since for the ease of archivation and reporting, the insurance company needs to relate all this history to a single year.

Once we choose this year, we can reflect this information in our database:

```{r, echo=F}
tibble::tribble(
  ~AY, ~DY, ~Year, ~Amount, ~Note,
  NA,   NA,  2021,  0,       "Policy Start",
  2022, 1,   2022,  0,       "Accident (Driver A reports to insurer)",
  2022, 2,   2023,  0,       "Claim submitted by Person B",
  2022, 2,   2023,  10000,   "Car",
  2022, 2,   2023,  1000,    "Wage compensation for 2022/12",
  2022, 2,   2023,  3000,    "Surgery1",
  2022, 2,   2023,  4000,    "Surgery2",
  2022, 2,   2023,  12000,   "Wage compensation for 2023/01-2023/12",
  2022, 3,   2024,  12000,   "Wage compensation for 2024/01-2024/12",
  2022, 4,   2025,  2400,    "Partial disability compensation",
  2022, 5,   2026,  2400,    "Partial disability compensation",
)
```

We can aggregate this information for each combination of AY and DY, then display in a wide (pivoted) format like this:

```{r, echo=F}
tibble::tribble(
  ~AY, ~DY1, ~DY2, ~DY3, ~DY4, ~DY5,
  2022, 0, 30000, 42000, 44400, 46800
)
```

However, since we're an insurance company, *we're interested in looking at the whole portfolio* (multiple policies and therefore multiple claims). Let's assume we had 10 similar claims that originated in 2022 (accident year):

```{r, echo=F}
tibble::tribble(
  ~AY, ~DY1, ~DY2, ~DY3, ~DY4, ~DY5,
  2022, 0, 300000, 420000, 444000, 468000
)
```

Working with a bigger portfolio also means, that *some claims usually happen every year*, not just in 2022. So a *full picture of our portfolio* could look something like this. Note that as we progress in time (AY), we have less and less information on how the losses will look like in the following development years (DY).

This structure is shaped like a triangle, where the rows represent the years of origin (policy, accident or reporting year) and the columns represent how long it took until the information was recognized.

```{r, echo=F}
tibble::tribble(
  ~AY, ~DY1, ~DY2, ~DY3, ~DY4, ~DY5,
  2022, 0, 300000, 420000, 444000, 468000,
  2023, 10000, 320000, 410000, 454000, NA,
  2024, 12000, 350000, 430000, NA, NA,
  2025, 6000, 305000, NA, NA, NA,
  2026, 9000, NA, NA, NA, NA
)
```

### What do we do with this?

Well, we want to take the advantage of the history that we have collected and predict what is likely to happen with the latest year, that we have just got to know about. It is basically a variation of the Linear Regression, where you learn from past patterns and apply them to predict future.

To get what we are talking about, it is worth visualizing this a bit. Use the ChainLadder package and explore the triangle object from example there. The difference between the predicted ultimate losses and current paid claims is our reserve (IBNR — Incurred But Not Reported/Recorded).
