---
title: "Pre-Model Analysis and data preparation "
output: html_notebook
---

### Exercise 1
> go to `loss_analysis.R` and make a few data checks

# Policy vs. Claims data
> What data do we need for modelling? What are we going to model actually?
> This is crucial question for the business problem you are trying to solve.

### Exercise 2
> Go to Exercise 2 and try to think about what could be a good 'target'/ event 
> you are trying to predict from data you have.

## Modeling data level
> Next question you should ask yourself before you jump directly into modeling is 
> How granular you need data to be prepared. 
> What makes sense to model in our case? Is it quality of the potential/future policy?
> Is it claim amount or something completely different?

Claims data are at *claims* level.
```{r}
dt_Claims <- read.csv("../../data/claims.csv")
dt_Claims %>% nrow()
dt_Claims %>% distinct(NrClaim) %>% nrow()
dt_Claims %>% distinct(NrPolicy) %>% nrow() 
```
Policy data are at *policy* level.
```{r}
dt_Policy <- read.csv("../../data/policy_history.csv")
dt_Policy %>% nrow()
dt_Policy %>% distinct(NrClaim) %>% nrow()
dt_Policy %>% distinct(NrPolicy) %>% nrow() 
```

## tie cisla su divne, nema to byt naopak?

## Target creation

