---
title: "Tidyverse"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

* collection of very usefull R packages for data wrangling, visualization etc. 
* we will use mainly two, `dplyr` for data preparation and `ggplot2` for visualizations
* package cheatsheets and simple example, we will not go through details, learning by doing
* link to literature, vignettes and nice examples

-------------------------------

## dplyr - data transformation package
`dplyr` is well known package for data tranformation especially in recent R community.

Using concept [grammar of data transformation](https://rstudio-pubs-static.s3.amazonaws.com/153547_3a8e24d4ceeb4868a5f22e70641e8165.html#/) providing consistent set of verbs that help you solve the most common data manipulation challenges.

You can solve ~85% of data manipulation issues using 6 verbs:

-   `mutate()` adds new variables that are functions of existing variables
-   `select()` picks variables based on their names.
-   `filter()` picks cases based on their values.
-   `summarise()` reduces multiple values down to a single summary.
-   `arrange()` changes the ordering of the rows.

These all combine naturally with `group_by()` which allows you to perform any operation "by group". You can learn more about them in `vignette("dplyr")`. As well as these single-table verbs, dplyr also provides a variety of two-table verbs, which you can learn about in `vignette("two-table")`.

### Exercise - dplyr
```{r message = FALSE}
library(dplyr)
```

#### Piping
```{r}
abs(min(c(1, 4, 2, 8, -5)))
```

vs.
```{r}
c(1, 4, 2, 8, -5) %>% min %>% abs
```
#### 6 basic dplyr verbs
```{r}
dt_KPI <- read.csv("../../data/lesson2_KPI.csv")

dt_KPI %>% head
```

```{r}
dt_KPI %>% 
  mutate(Combined = Losses + Expenses) %>% head
```

```{r}
dt_KPI %>% 
  select(Region, Losses) %>% head
```

```{r}
dt_KPI %>% 
  filter(Business == "Criminals") %>% head
```

```{r}
dt_KPI %>% 
  arrange(Losses, desc(Premium)) %>% head
```

```{r}
dt_KPI %>% 
  summarise(Loss_avg = mean(Losses)) %>% head
```

```{r}
dt_KPI %>% 
  group_by(Region) %>% 
  summarise(Loss_avg = mean(Losses)) %>% head
```

## ggplot2 - data visualization package
>“The simple graph has brought more information to the data analyst’s mind than any other device.” — John Tukey

Althoug R has a lot of systems of providing graphics, `ggplot2` is one of the most elegant and most versatile system for visualiaztion in R. ggplot2 implements the [grammar of graphics](http://vita.had.co.nz/papers/layered-grammar.html), a coherent system for describing and building graphs.

### First Plot - Scatter
```{r}
library(ggplot2)

ggplot(data = dt_KPI) +
  geom_point(mapping = aes(x = Premium, y = Expenses))
```

### Aesthetics and layers
```{r}
ggplot(data = dt_KPI) +
  geom_point(mapping = aes(x = Premium, y = Expenses, colour = Region))
```

### Geometric Objects and iheritance of aesthetics
```{r message = FALSE}
ggplot(data = dt_KPI,
       mapping = aes(x = Premium, y = Expenses, colour = Region)
) +
geom_point() +
geom_smooth()
```

## Other Material, Literature and Credits
[R for Data Science - Data Transformation](http://r4ds.had.co.nz/transform.html#introduction-2)

[R for Data Science - Data Visualization](http://r4ds.had.co.nz/data-visualisation.html)


[dplyr cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf)

[ggplot2 cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf)
