---
title: "Exercise - ggplot2"
output: html_notebook
---

[Lesson 2 on Github](https://github.com/ex-man/GeneralInsurance_Class/blob/Class2023/Lessons/Lesson2/README.md)

> Run following command in R console to link to `ggplot2` library:

```{r}
# install.packages('ggplot2')
library(ggplot2)
library(dplyr)
```

> Load data if you have not done so yet:

```{r}
dt_KPI <- read.csv("./Data/lesson2_KPI.csv")
```

### First Plot - Scatter
> Now let's try to show our data on simple scatter plot

```{r}
ggplot(data = dt_KPI, aes(x = Premium, y = Expenses)) +
  geom_point()

dt_KPI %>%
  ggplot(aes(x = Premium, y = Expenses)) +
  geom_point()

dt_KPI %>%
  ggplot(aes(x = Region, y = Premium)) +
  geom_col()
```

> We used `ggplot` function specifying *data* set to use, `aes` function to specify *aestetics* (visualisation) and *geoms* function `geom_point` to specify how data should be represemeted.  
> Note that `+` serves as sign to add geom layer to a chart. We will use this technique in future steps as well.  


### Aesthetics and layers
> In previous example we set *data* and *aesthetics* on global level, setting axes in *geoms* to *Premium* and *Expenses*. We will do the same now but specifying it for geom `geom_points` only. This is usefull when you want to show multiple measures on same chart having their own axes.  
> In addition we will also specify colour of dots varying colour by *Region*.  

```{r}
ggplot(data = dt_KPI) +
  geom_point(mapping = aes(x = Premium, y = Expenses, colour = Region))
```


### Geometric Objects and iheritance of aesthetics
> Now let's add new *geom* for smoothed regresion line. We will do so by adding new *geom* using `+` sign and function `geom_smooth`.

```{r}
ggplot(data = dt_KPI) +
  geom_point(mapping = aes(x = Premium, y = Expenses, colour = Region)) +
  geom_smooth(mapping = aes(x = Premium, y = Expenses)) +
  geom_smooth(mapping = aes(x = Premium, y = Expenses, color = Region))
```

> Note that there is one regresion line for whole dataset. Also note that we can specify `x` and `y` for each *geom* separatelly.  
> This happens when we specify `aes` on global (`ggplot`) level:
ggplot(data = dt_KPI,

```{r}
ggplot(
    data = dt_KPI,
    mapping = aes(x = Premium, y = Expenses, colour = Region)
  ) +
  geom_point() +
  geom_smooth()
```

> It produced chart with dots and smoothed line where regresion is done on data by *Region*.  
> This feature of `ggplot` is called *inheritance*. You can do lot's using it.

> Visit https://github.com/rstudio/cheatsheets/blob/main/data-visualization.pdf for more information about ggplot2